### 2015-06-05 LOI/inc2.tex
Initial edits -- a work in progress

\underline{\textbf{Objectives:}} Exascale scientific discovery will
%introduce more complex hardware, and many simulations will 
be bottlenecked 
without sufficient new research into managing and storing the large data which will be
produced during the simulation, and analzyed for months after the simulations.
Our goal is to investigate a unique approach to addressing this challenge in
a multi-tier storage platform, expediting insights into mission critical
scientific processes. 
% the multi-tier challenges that are faced by scientists in creating and
% managing and storing their data  to expedite insights xsinto
% mission critical scientific processes in exascale compuxsting.
%
We propose a research program that aims to explore a hiearchical
organization and storage infstraturure which will facilitiate how data can
be written to and read from a complex heterogenous storage hierarchy
efficiently.  
%
We will apply a technique which will re-organize and possibly
reduce the total amount of information based on user intentions along with
direct feedback from the storage and I/O system to allow data to be
optimized not just for output, but also for many of the complex reading
patterns. Since data analysis and visualization will be performed in both an
in situ and post processing environment, we will continue to use our I/O and
storage abstractions to help enable applications to take full advantage of
the hardware and softare environments. Because data volumes are strictly
outstripping available bandwidths, compression will be required. Rather than
focusing solely on lossless compression, we will incorporate lossy
compression with the ability to annotate data allowing optimized algorithm
selection based on the numerical methods employed in the source
simulation. We plan to epxlore this tradeoff of storage capacity and
bandwidth for computation to determine how to offer the user options that
can meet their needs with desireable performance characteristics.

In order to explore the many research challenges in this
domain, we will build upon the success of our middleware system, ADIOS, our
multi-tier storage system, Sirocco, and our distributed storage system,
Ceph.  We will also leverage data movement and staging technologies, Data
Spaces and the Common Communication Interface (CCI), as part of a software
stack used successfully on current leadership class machines. Our experience
with this stack will guide us in understanding the new directions for
research in storage systems and I/O for exascale and beyond systems.

@@

% which means that we will
% also support additional access modes, beyond the traditional
% Checkpoint/Restart methods, and understand that all data has errors
% associated with it and we will ensure that these become encoded to allow for
% additional optimizations. 
For example, all scientific simulations contain
approximations, and all measurements from obervations and experiments also
contain errors. These errors can help guide how information can be presented
to users, by allowing us to ask for information within a given accuracy
bound, and also allow us to offer a mode for recomputation vs. data storage
and reterival.

@@

Our approach will allow users to ``plug-in'' techniques to classify data, not as bytes but as motifs, where
we can understand relationships between objects (into data models) and relationships of data in variables.
This will allow us to adapt data from a variable into the various storage tiers. Rather than focus on simple data
compression, we will incoroarte a multitude of techniques to reduce data on the faster storage tiers, and keep the less
reduced information on the lower tiers.

We wish to support additional data access modes as well. First, within a given
accuracy bounds, offer a mode for recomputation from data stored in a ``fast''
tier to reduce data latencies. Second, exploratory analysis operations
frequently entail an overall data set view followed by targeted data
exploration based on identified features. We will offer support for a
configurable data access mode to support these sorts of analysis accesses. An
``overview'' access mode that gives a quick, approximate within error bounds,
data view that can guide feature selection offering rapid coarse-grained data
exploration without requiring loading the complete, detailed data set from
storage. Based on the granularity requested, the accuracy and size of the data
returned can be adjusted. At the most extreme setting, the original data can
be retrieved at the time cost of moving the potentially huge data quantity.
Third, to ensure available storage for subsequent operations, we will offer
automatic data migration based on user annotations for required data lifetimes
using monitoring and learning techniques. Unlike existing approaches, this will
be tempered both by the user annotations and through learned access patterns.
While past access patterns may not indicate future access because the
simulation run purpose may have changed, we are focused on scalability where
runs are subsequently larger as the simulation prepares for a capability run.
By learning from the output and access patterns during this run sequence, we
can accurate decide how to place and organize data for the critical capability
runs. Fourth, we anticipiate storing mutliple data copies, each compressed in
different ways according to the underlying media, some of these copies will
disappear based on storage pressures, but data persistence will be maintained
according to user specifications. Assuming a relatively low latency cache layer
before a tape system, we can offer exporatory data access reserving pulling
data from tape to just the data required. This will save scientists time and
make data stored on tape usable without long delays.

@@

The success of this project will  provide insights into how to build middleware and storage layers which can interact
well with each other, and take user-provided hints. Today data is reduced by application scientist who have 
limited information on what the storage layer can provide. They often make compromises based on this limited knowledge
and either tune their output for writing or reading performance. This data then gets moved to other locations, and much of
the tuning is lost when the data is read back during their post processing. Furthermore, there is a limited set of operations
which users will be able to stage to other staging nodes for real-tiime-reduction and visualization.

>>>>

\underline{\textbf{Objectives:}} Exascale scientific discovery will
%introduce more complex hardware, and many simulations will 
be severely bottlenecked without sufficient new 
 research into managing and storing the large amounts of 
data that will be produced during the simulation, and analyzed for months after the simulations.
%Our goal in this project is to investigate a unique approach to addressing this data management 
%and storage challenge for multi-tier exascale storage architectures, thereby expediting insights into mission critical
%scientific processes. 
Our goal in this project is to address this data management and storage challenge. Specifically we plan
 to research and develop a unique application-aware hierarchy data organization that can effectively map 
 to  multi-tier exascale storage architectures, and can facilitate how data can be efficiently written to and 
 read from a complex heterogeneous storage hierarchy  
 thereby expediting insights into mission critical scientific processes. 
% the multi-tier challenges that are faced by scientists in creating and
% managing and storing their data  to expedite insights xsinto
% mission critical scientific processes in exascale compuxsting.
%
%Specifically, we propose a research program that aims to explore a hierarchical
%organization and storage infrastructure that will facilitate how data can
%be written to and read from a complex heterogenous storage hierarchy
%efficiently.  
%
Our approach is to explicitly and implicitly capture user/application intentions 
and use this information, along with direct feedback from the storage and I/O 
system, to dynamically structure, organize, process and map the data so that 
it can most effectively support complex I/O patterns. Furthermore, we will do 
this with an awareness of data usage semantics and with an understanding 
of current and future hardware and software technologies. 
%re-organize and possibly reduce the total amount of information 
%Since data analysis and visualization will be performed in both an
%in situ and post processing environment, we will continue to use our I/O and
%storage abstractions to help enable applications to take full advantage of
%the hardware and software environments. 
For example, as data volumes are quickly outgrowing available bandwidths, 
compression is becoming essential. Rather than focusing solely on lossless 
compression, we will incorporate lossy compression with the ability to annotate 
data allowing optimized algorithm selection based on the numerical methods 
employed in the source simulation. We plan to explore this tradeoff of storage 
capacity and bandwidth for computation to determine how to offer the user 
options that can meet their needs with desirable performance characteristics.

In order to explore the many research challenges in this domain, we will build 
upon the success of our middleware system, ADIOS, our multi-tier storage 
system, Sirocco, and our distributed storage system, Ceph.  We will also leverage 
data staging and data movement technologies, DataSpaces and the Common 
Communication Interface (CCI), as part of a software stack. 
%used successfully on current leadership class machines. 
%Our experience with this stack will guide us in understanding the new directions for
%research in storage systems and I/O for exascale and beyond systems.

@@

% which means that we will
% also support additional access modes, beyond the traditional
% Checkpoint/Restart methods, and understand that all data has errors
% associated with it and we will ensure that these become encoded to allow for
% additional optimizations. 
For example, all scientific simulations contain
approximations, and all measurements from observations and experiments also
contain errors. These errors can help guide how information can be presented
to users, by allowing us to ask for information within a given accuracy
bound, and also allow us to offer a mode for re-computation vs. data storage
and reterival.

@@

Our approach will allow users to ``plug-in'' techniques to classify data, not as bytes but as motifs, where
we can understand relationships between objects (into data models) and relationships of data in variables.
This will allow us to adapt data from a variable into the various storage tiers. Rather than focus on simple data
compression, we will incorporate a multitude of techniques to reduce data on the faster storage tiers, and keep the less
reduced information on the lower tiers.

We wish to support additional data access modes as well. First, within a given
accuracy bounds, offer a mode for re-computation from data stored in a ``fast''
tier to reduce data latencies. Second, exploratory analysis operations
frequently entail an overall data set view followed by targeted data
exploration based on identified features. We will offer support for a
configurable data access mode to support these sorts of analysis accesses. An
``overview'' access mode that gives a quick, approximate within error bounds,
data view that can guide feature selection offering rapid coarse-grained data
exploration without requiring loading the complete, detailed data set from
storage. Based on the granularity requested, the accuracy and size of the data
returned can be adjusted. At the most extreme setting, the original data can
be retrieved at the time cost of moving the potentially huge data quantity.
Third, to ensure available storage for subsequent operations, we will offer
automatic data migration based on user annotations for required data lifetimes
using monitoring and learning techniques. Unlike existing approaches, this will
be tempered both by the user annotations and through learned access patterns.
While past access patterns may not indicate future access because the
simulation run purpose may have changed, we are focused on scalability where
runs are subsequently larger as the simulation prepares for a capability run.
By learning from the output and access patterns during this run sequence, we
can accurate decide how to place and organize data for the critical capability
runs. Fourth, we anticipate storing multiple data copies, each compressed in
different ways according to the underlying media, some of these copies will
disappear based on storage pressures, but data persistence will be maintained
according to user specifications. Assuming a relatively low latency cache layer
before a tape system, we can offer exploratory data access reserving pulling
data from tape to just the data required. This will save scientists time and
make data stored on tape usable without long delays.

@@

The success of this project will  provide insights into how to build middleware and storage layers which can interact
well with each other, and take user-provided hints. Today data is reduced by application scientist who have 
limited information on what the storage layer can provide. They often make compromises based on this limited knowledge
and either tune their output for writing or reading performance. This data then gets moved to other locations, and much of
the tuning is lost when the data is read back during their post processing. Furthermore, there is a limited set of operations
which users will be able to stage to other staging nodes for real-time-reduction and visualization.
